### 2021-06-17 paper/abstract.tex
No obvious stuff left, at least nothing obvious at this time of the night.

\noindent
% Abstracts of Research Articles and Reports should explain to the general reader why
% the research was done, what was found and why the results are important. 
% 
% They should start with some brief BACKGROUND information: a sentence giving a broad
% introduction to the field comprehensible to the general reader, and then a sentence of
% more detailed background specific to your study. 
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially,
these focused on non-pharmaceutical interventions; more recently, vaccinations and
large-scale rapid testing have started to play a major role. The objective of this study
is to explain the quantitative effects of these policies on determining the course of
the pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. In order to do so, the study develops an agent-based simulation
model that is estimated on data for the second and the third wave of the CoViD-19
pandemic in Germany. The paper finds that during a period where vaccination rates rose
from 5\% to 40\%, large-scale rapid testing had the largest effect on reducing infection
numbers. 
% The final sentence should outline the main CONCLUSIONS of the study, in terms that
% will be comprehensible to all our readers. The Abstract is distinct from the main body
% of the text, and thus should not be the only source of background information critical
% to understanding the manuscript. Please do not include citations or abbreviations in
% the Abstract. 
Frequent, large-scale rapid testing should remain part of strategies to contain
CoViD-19; it can substitute for many non-pharmaceutical interventions that come at a
much larger cost to individuals, society, and the economy.

>>>>

\noindent
% Abstracts of Research Articles and Reports should explain to the general reader why
% the research was done, what was found and why the results are important. 
% 
% They should start with some brief BACKGROUND information: a sentence giving a broad
% introduction to the field comprehensible to the general reader, and then a sentence of
% more detailed background specific to your study. 
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially,
these focused on non-pharmaceutical interventions; more recently, vaccinations and
large-scale rapid testing have started to play a major role. The objective of this study
is to explain the quantitative effects of these policies on determining the course of
the pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. To do so, the study develops an agent-based simulation
model, which is estimated using data for the second and the third wave of the CoViD-19
pandemic in Germany. The paper finds that during a period where vaccination rates rose
from 5\% to 40\%, large-scale rapid testing had the largest effect on reducing infection
numbers. 
% The final sentence should outline the main CONCLUSIONS of the study, in terms that
% will be comprehensible to all our readers. The Abstract is distinct from the main body
% of the text, and thus should not be the only source of background information critical
% to understanding the manuscript. Please do not include citations or abbreviations in
% the Abstract. 
Frequent large-scale rapid testing should remain part of strategies to contain
CoViD-19; it can substitute for many non-pharmaceutical interventions that come at a
much larger cost to individuals, society, and the economy.

### 2021-06-17 paper/model/course_of_disease.tex
Got started re-writing with Radost's comments on the side.
No obvious stuff left, at least nothing obvious at this time of the night.

The following medical parameters describing the progression of the disease are taken from
systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
progresses in the way depicted in Figure~\ref{fig:disease_progression}.

@@

We allow the progression of the disease to be stochastic in two ways: Firstly, state
changes only occur with a certain probability (e.g. only a fraction of infected
individuals develops symptoms). Secondly, the number of periods for which an individual
remains in a state is drawn randomly. The parameters that govern these processes are
taken from the literature\footnotemark and age-dependent.

\footnotetext{ Detailed information on the calibration of the disease parameters is
    available as part of our
    \href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
    documentation}. }

>>>>

The following medical parameters describing the progression of the disease are taken from
systematic reviews (e.g. \citet{He2020}). After an infection occurs, the disease
progresses in the way depicted in Figure~\ref{fig:model_disease_progression}.

@@

We allow the progression of the disease to be stochastic in two ways: Firstly, state
changes only occur with a certain probability (e.g. only a fraction of infected
individuals develops symptoms). Secondly, the number of periods for which an individual
remains in a state is drawn randomly. The parameters that govern these processes are
taken from the literature \footnote{ Detailed information on the calibration of the
disease parameters is available as part of our
\href{https://sid-dev.readthedocs.io/en/latest/reference_guides/epi_params.html}{online
documentation}.} and age-dependent.

### 2021-06-17 paper/model/policies.tex
No obvious stuff left, at least nothing obvious at this time of the night.

\subsection{Reducing Numbers of Contacts Through Policies}
\label{sec:policies}

@@

On an abstract level, a policy is a functions that modifies the number of contacts of
one contact type. For example, school closures simply set all school contacts to zero. A
lockdown where only essential workers are allowed to work means that approximately two
thirds of the working population have zero work contacts and the rest has the same
number of contacts as before.

>>>>

\subsection{Reducing Numbers of Contacts via NPIs}
\label{sec:policies}

@@

On an abstract level, a policy is a functions that modifies the number of contacts of
one contact type. For example, school closures simply set all school contacts to zero. A
lockdown where only essential workers are allowed to work means that approximately two
thirds of the working population have zero work contacts and the rest has the same
number of contacts as before.\comment[id=HM]{Mention here or in results that school
closures $\neq$ vacations because of additional contacts}

### 2021-06-17 paper/report.tex
No obvious stuff left, at least nothing obvious at this time of the night.

This paper develops a quantitative model incorporating these factors simultaneously. The
framework allows to combine a wide variety of data and mechanisms in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany and show that rapid testing had the largest impact on the reduction in
infections by almost 80\% during the month of May 2021. We conclude that rapid tests
have a large role to play for at least as long as vaccinations have not been offered to
an entire population.

@@

% The thresholds fall over time and are calibrated to survey data where possible (see
% Supplementary Materials~\ref{subsec:rapid_test_demand}).

% The thresholds are higher for tests in a private setting than for tests at the
% workplace until mid April and then compliance because of private reasons becomes
% higher than compliance at the workplace.

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests.

The free parameters are the infection probabilities by contact category, one hygiene
multiplier for school and work contacts after November 2nd 2020, approx. ten parameters
that govern the reduction of other contacts\comment[id=K]{Must be updated when the
estimation is finished.}, the number of extra contacts during holidays, the share of
detected cases around holidays and the fade in speed of the rapid tests as well as one
parameter that governs the import of B.1.1.7 cases in January 2021. For details, see
Supplementary Material~\ref{subsec:estimated_params},~\ref{subsec:rapid_test_demand}.

We estimate these parameters such that the model matches the following empirical
quantities: 1. official case numbers for each age group and region, 2. deaths, 3. share
of B.1.1.7. \citep[See][for the general method]{McFadden1989}. However, only a fraction
of infections actually enter the official case numbers. We thus model official cases as
depicted in Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the
share of detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As
the share of asymptomatic individuals varies by age group, this gives us age-specific
shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases
by age group over time in our model). Our estimates suggest that only 18 to 28\% of cases
are detected for school age children while the rate is 30 to 55\% for those above age
80.\comment[id=K]{Must be verified once results are final.} Confirmation of a positive
rapid test result is another reason for PCR tests, once these become available.

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. For legibility, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value 1 represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of 10. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November, before rising again in December, which prompted the most stringent lockdown to
this date. Schools and daycare centers were closed again, so were customer-facing
businesses except for grocery and drug stores. From the peak of the second wave just
before Christmas until the trough in mid-February, newly detected cases decreased by
almost three quarters. The third wave in the spring of 2021 is associated with the
B.1.1.7 strain, which became dominant in March. See Figure~\ref{fig:share_b117}. In
early March, some NPIs were being relaxed; e.g., hairdressers and home improvement
stores were allowed to open again to the public. There were many changes in details of
regulations afterwards, but they did not change the stringency index.

@@

@@

        \caption{{Tests and vaccinations}}
        \label{fig:antigen_tests_vaccinations}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering

        \caption{Fraction of B.1.1.7 strain}
        \label{fig:share_b117}
    \end{subfigure}

@@

By this time, the set of policy instruments had become much more diverse. Around the
turn of the year, the first people were vaccinated with a focus on older age groups and
medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just
over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
started to replace regular PCR tests for staff in many medical and nursing facilities.
These had to be administered by medical doctors or in pharmacies. At-home tests approved
by authorities became available in mid-March, rapid test centers were opened and one
test per person and week was made available free of charge. Depending on the state,
customers were only allowed to enter certain stores with a recent negative rapid test
result. These developments are characteristic of many countries: The initial focus on
NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
acceptance and use of rapid tests. At broadly similar points in time, novel strains of
the virus have started to pose additional challenges.

We draw simulated samples of agents from the distribution of recorded infections in
September 2020 and use the model to predict recorded infection rates until the end of May
2021.\comment[id=J]{This needs more explanation. I will rewrite this.} See Supplementary
Materials~\ref{sec:data_and_parameters} and \ref{sec:model} for detailed descriptions of
the data and the model, respectively. The blue line in Figure~\ref{fig:aggregated_fit}
shows our model's predictions are very close to officially recorded cases in the
aggregate. This is also true for infections by age and geographical region, which are
shown in the supplementary materials (Figures~\ref{fig:age_group_fit} and
\ref{fig:state_fit}, respectively).

The effects of various mechanisms can be disentangled due to the distinct temporal
variation in the drivers of the pandemic. Next to the stringency index, the three lines
in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
increased hygiene regulations, and seasonality evolved since early September for each of
the three broad contact networks. For example, a value of 0.75 for the work multiplier
means that if the environment was the same as in September (levels of infection rates,
no rapid tests or vaccinations, only the wildtype virus present), infections at the
workplace would be reduced by 25\%. The lines show the product of the effect of contact
reductions, increased hygiene regulations, and seasonality. Two aspects are particularly
interesting. First, all lines broadly follow the stringency index and they would do so
even more if we left out seasonality and school vacations (roughly the last two weeks of
October, two weeks each around Christmas and Easter, and some days in late May). Second,
the most stringent regulations are associated with the period of strong decreases in new
infections between late December 2020 and mid-February 2021. The reversal of the trend
is associated with he spread of the B.1.1.7 variant. The steep drop in recorded cases
during May 2021 is associated with at least weekly rapid tests to around 42~percent of
the population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
impacting a fall in the relative infectiousness of contacts outside of work and school.

In order to better understand the contributions  of rapid tests, vaccinations, and of
seasonality on the evolution of infections in 2021, Figure~\ref{fig:2021_scenarios_broad}
considers various scenarios. NPIs are always the same as in the baseline scenario.
Figure~\ref{fig:2021_scenarios_recorded} shows the model fit (the blue line, same as in
Figure~\ref{fig:aggregated_fit}), a scenario without any of the three factors (red line),
and three scenarios turning these factors off one by one.
Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total infections in the
model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley values
\citep{Shapley2016} to decompose the difference in total infections between the scenario
without any of the three factors and our main specification.\comment[id=J]{I would not
mention the word "coalition" in the notes to the figure on the Shapley values.
Combination with other channels is better. I think I would prefer to have a short section
on shapley values in the appendix (why use them, how do they work) and just link to that
in the figure note}

@@

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute about 15\% to
the cumulative difference between scenarios. Reasons for this are the slow start---it
took until March~24th until 10\% of the population had received their first vaccination,
the 20\% mark was reached on April 19th---and the focus on older individuals. These
groups contribute less to the spread of the disease than others due to a lower number of
contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. It is important to
note that the initial focus of the campaign was to prevent deaths and severe disease;
the case fatality was rate considerably lower during the third wave when compared to the
second (4.4\% between October and February and 1.4\% between March and June). By the end
of our study period, when first-dose vaccination rates reached around 40\% of the
population, the numbers of new cases would have started to decline.

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace. Nevertheless, we estimate it to be a quantitatively important factor
determining the evolution of the pandemic, explaining most of the early changes and
almost 40\% of the cumulative difference by the end of May.

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that infections are detected, which would otherwise
remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this means that
until late April, recorded cases are higher than in the scenario where none of the three
mechanisms are turned on. Compared to the scenario with vaccinations only, this point is
reached around mid-May and it would be June for the comparison with the seasonality-only
scenario. The effect on total cases, however, is visible immediately. Despite the fact
that only a small fraction of the population performed weekly rapid tests in
March,\footnote{4\% to 6\% have done a rapid test within the last seven days in the first
week of March. 8\% in the model have ever done a rapid test at that point; 27\% of
respondents to the COSMO study reported in early March that they \textit{ever} did a
rapid test)} new infections on April 1 would be reduced by 53\% relative to the
scenario without vaccinations, rapid tests, or seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only (purple line in
Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the three
channels for rapid tests. Tests for pupils have the smallest effect, which is largely
explained by the relatively small number of students (Only 18\% of our population are in
the education sector (pupils, teachers etc.) versus  46\% are workers outside the
education sector) and the fact that schools did not operate at full capacity during our
period of study. Almost 40\% come from tests at the workplace. Despite the fact that
rapid tests for private reasons are phased in only late, they make up for more than half
of the total effect.\comment[id=HM]{How many tests are actually performed along each of
the three dimensions? Could this be a quantity story rather than the directed testing
story I am making up here as wishful thinking?}\comment[id=K]{See subsection
\ref{subsec:appendix_rapid_tests}.}

The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
Kretzschmar2020}. This is also the reason for why the number of tests performed fall at
the very end of our study period. Falling infection rates mean that there are fewer
events triggering private test demand.\comment[id=HM]{Positive rates along the three
channels might be nice here, if previous comment is not a quant story}\comment[id=K]{See
figure~\ref{fig:share_of_rapid_tests_by_reason_by_infected}} \comment[id=HM]{I think
there is a story in here about risk conditional on past exposure vs. prospective risk in
gatherings, but it is too late to work through it. We might just want to calculate the
probabilities.}\comment[id=K]{I don't understand which probabilities you mean.}

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December on. Some states started opening
them gradually in late February, but usual operation was not back until the beginning of
June. Figure~\ref{fig:school_scenarios} shows the effects of different policies regarding
school starting at Easter, at which point rapid tests had become widely available. We
estimate the realized scenario to have essentially the same effect as a situation with
closed schools. Under fully opened schools with mandatory tests, total infections would
have been 7\% higher; this number rises to 20\% without tests. These effect sizes are
broadly in line with empirical studies, e.g., \citet{Vlachos2021}\comment[id=HM]{Need a
back-of the envelope calculation} To use another metric, the effective weekly
reproduction number differs by $0.017$ and $0.045$, respectively. In light of the large
negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021} \comment[id=HM]{need to cite a couple of poor learning outcomes / mental
health papers. There must be well-published ones}\comment[id=K]{I googled and found two
that I think are at least not terrible but I also didn't find anything really well
published}---and in particular on those with low socio-economic status---these results in
conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when no test
strategy is available, trade-offs may well be different.

@@

Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of employees
receiving tests, testing at the workplace has larger effects than changing requirements
to work from home. Whether the share of workers working at the usual workplace is
reduced or increased by ten percent hardly has an effect on infection rates. Making
testing mandatory twice a week---assuming independent compliance by employers and workers
or 95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
employers to the level of March would have increased them by the same amount.

Despite these large effects, the results on testing likely understate the benefits.
Disadvantaged groups of the population are less likely to be reached by testing
campaigns;\comment[id=HM]{seems obvious, but I did not find anything. COSMO data? We
should get the raw data, btw.} at the same time they have a higher risk to contract
CoViD-19.\citet{KochInstitut2021a} Mandatory tests at school and at the workplace will
extend more into these groups. The same goes for individuals who exhibit a low level of
compliance with CoViD-19-related regulations.

Our analysis\comment[id=J]{The conclusion should be slightly longer and more separated
from the rest} has shown that in the transition path to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A day of the winter lockdown in Germany is estimated to have
cost 6.25~billion~Euros \citep{Dorn2020b}.\comment[id=K]{This is a prediction by the ifo
Instute back from March 2020 assuming a 2 month lockdown. I did not find anything
citable that is more recent. I did find
\href{https://www.merkur.de/wirtschaft/lockdown-coronavirus-kosten-deutschland-milliarden-euro-wirtschaft-krise-zr-90231189.html}{A
quote from ifo's Timo Wollmershäuser: „Wir liegen zwischen zwei bis zweieinhalb
Milliarden Euro pro Woche“} for the winter lockdown} In early June 2021, retail prices
for self-administered rapid tests are frequently less than
1~Euro.\comment[id=HM]{Cheapest on \href{https://amazon.de}{Amazon} on June 11th, 2021.
--- DM 95ct}

>>>>

This paper develops a quantitative model incorporating these factors simultaneously. The
framework allows to combine a wide variety of data and mechanisms in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany, where new infections fell by almost 80\% during the month of May 2021. Our
analysis shows that frequent and large-scale rapid testing caused the bulk of this
decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that it
should have a large role for at least as long as vaccinations have not been offered to
an entire population.

@@

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and estimate the model's many
parameters.\footnote{See section~\ref{sec:data_and_parameters} of the supplementary
materials for an overview.} Contact diaries yield pre-pandemic distributions of contacts
for different contact types and their assortativity by age group. Mobility data is used
to model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method, also described in
Supplementary Material \ref{sec:model}]{McFadden1989}.\comment[id=HM]{Make reference
more specific once Appendix has new structure}. We keep the number of free parameters
low in order to avoid overfitting. In our application, these features include  official
case numbers for each age group and region, deaths, and the share of the B.1.1.7 strain.

The main issue with official case numbers is that they will contain only a fraction of
all infections. We thus model official cases as depicted in
Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As the share
of asymptomatic individuals varies by age group, this gives us age-specific shares (see
Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age
group over time in our model). Our estimates suggest that---in the absence of rapid
testing---the detection rate is two to three times higher for individuals above age 80
compared to school age children.\comment[id=K]{Must be verified once results are final.}
Once rapid test become available, confirmation of a positive result is another reason
leading to positive PCR tests.

The model is applied to the second and third wave of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
never use it directly. For legibility reasons, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value 1 represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of 10. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November before rising again in December, prompting the most stringent lockdown to this
date. Schools and daycare centers were closed again, so were customer-facing businesses
except for grocery and drug stores. From the peak of the second wave just before
Christmas until the trough in mid-February, newly detected cases decreased by almost
three quarters. The third wave in the spring of 2021 is associated with the B.1.1.7
strain, which became dominant in March (Figure~\ref{fig:share_b117}). In early March,
some NPIs were being relaxed; e.g., hairdressers and home improvement stores were
allowed to open again to the public. There were many changes in details of regulations
afterwards, but they did not change the stringency index.

@@

        \caption{Fraction of B.1.1.7 strain}
        \label{fig:share_b117}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.475\textwidth}
        \centering

@@

        \caption{{Tests and vaccinations}}
        \label{fig:antigen_tests_vaccinations}
    \end{subfigure}

@@

By this time, the set of policy instruments had become much more diverse. Around the
turn of the year, the first people were vaccinated with a focus on older age groups and
medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}). By the end of May, just
over 40\% had received at least one dose of a vaccine. Around the same time, rapid tests
started to replace regular PCR tests for staff in many medical and nursing facilities.
These had to be administered by medical doctors or in pharmacies. At-home tests approved
by authorities became available in mid-March, rapid test centers were opened and one
test per person and week was made available free of charge. Depending on the state,
customers were only allowed to enter certain stores with a recent negative rapid test
result. These developments are characteristic of many countries: The initial focus on
NPIs to slow the spread of the disease has been accompanied by vaccines and a growing
acceptance and use of rapid tests. At broadly similar points in time, novel strains of
the virus have started to pose additional challenges.

We draw simulated samples of agents from the population structure in September 2020 and
use the model to predict recorded infection rates until the end of May 2021. See
Supplementary Material~\ref{sub:initial_conditions} for a detailed description of this
procedure. The blue line in Figure~\ref{fig:aggregated_fit} shows our model's
predictions are very close to officially recorded cases in the aggregate. This is also
true for infections by age and geographical region, which are shown in the supplementary
materials (Figures~\ref{fig:age_group_fit} and \ref{fig:state_fit}, respectively).

The effects of various mechanisms can be disentangled due to the distinct temporal
variation in the drivers of the pandemic. Next to the stringency index, the three lines
in Figure~\ref{fig:stringency_infectious_contacts} summarize how contact reductions,
increased hygiene regulations, and seasonality evolved since early September for each of
the three broad contact networks. For example, a value of 0.75 for the work multiplier
means that if the environment was the same as in September (levels of infection rates,
no rapid tests or vaccinations, only the wildtype virus present), infections at the
workplace would be reduced by 25\%. The lines show the product of the effect of contact
reductions, increased hygiene regulations, and seasonality. Two aspects are particularly
interesting. First, all lines broadly follow the stringency index and they would do so
even more if we left out seasonality and school vacations (roughly the last two weeks of
October, two weeks each around Christmas and Easter, and some days in late May). Second,
the most stringent regulations are associated with the period of strong decreases in new
infections between late December 2020 and mid-February 2021. The reversal of the trend
is associated with the spread of the B.1.1.7 variant. The steep drop in recorded cases
during May 2021 is associated with at least weekly rapid tests to 42\% of the
population, a vaccination rate that rose from 28\% to 43\%, and mostly seasonality
impacting a fall in the relative infectiousness of contacts outside of work and school.

In order to better understand the contributions  of rapid tests, vaccinations, and
seasonality on the evolution of infections in 2021,
Figure~\ref{fig:2021_scenarios_broad} considers various scenarios. NPIs are always the
same as in the baseline scenario. Figure~\ref{fig:2021_scenarios_recorded} shows the
model fit (the blue line, same as in Figure~\ref{fig:aggregated_fit}), a scenario
without any of the three factors (red line), and three scenarios turning these factors
off one by one. Figure~\ref{fig:2021_scenarios_newly_infected} does the same for total
infections in the model. Figure~\ref{fig:2021_scenarios_decomposition} employs Shapley
values \citep{Shapley2016} to decompose the difference in total infections between the
scenario without any of the three factors and our main specification.

@@

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute about 15\% to
the cumulative difference between scenarios. Reasons for this are the slow start---it
took until March~24th until 10\% of the population had received their first vaccination,
the 20\% mark was reached on April 19th---and the focus on older individuals. These
groups contribute less to the spread of the disease than others due to a lower number of
contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
study period, when first-dose vaccination rates reached around 40\% of the population,
the numbers of new cases would have started to decline. It is important to note that the
initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
fatality was rate considerably lower during the third wave when compared to the second
(4.4\% between October and February and 1.4\% between March and the end of May).

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace \citep[this is in line with results in][]{Gavenciak2021}. Nevertheless,
we estimate seasonality to be a quantitatively important factor determining the
evolution of the pandemic, explaining most of the early changes and close to 40\% of the
cumulative difference by the end of May.

The largest effect---almost one half when considering the decompositions---comes from
rapid testing. Here, it is crucial to differentiate between recorded cases and actual
cases. Additional testing means that infections are recorded, which would otherwise
remain undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is
large. Until late April, recorded cases are higher in the scenario with rapid testing
alone when compared to the setting where none of the three mechanisms are turned on.
Compared to the scenario with vaccinations only, this point is reached around mid-May
and it would be June for the comparison with the seasonality-only scenario. The effect
on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only a small
fraction of the population performed weekly rapid tests in March,\footnote{4\% to 6\% of
respondents to the COSMO study report having done a rapid test during the first week of
March; 8\% in the model have ever done a rapid test at that point.}\comment[id=HM]{would
be great to have total tests performed in March, model vs. data is not important here.
Just first week of March looks too small} new infections on April 1 would be reduced by
53\% relative to the scenario without vaccinations, rapid tests, or seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only (purple line in
Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
three channels for rapid tests. Tests at schools have the smallest effect, which is
largely explained by schools not operating at full capacity during our period of study
and the relatively small number of students.\footnote{18\% of our population are in the
education sector (pupils, teachers etc.); 46\% are workers outside the education sector}
Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for
private reasons are phased in only late, they make up for more than half of the total
effect. The reason lies in the fact that a substantial share of these tests is driven by
an elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or
following up on a positive test of a household member. The latter is essentially a form
of contact tracing, which has been shown to be very effective \citep{Contreras2021,
Kretzschmar2020,FetzerGraeber2021}.

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December onwards. Some states started
opening them gradually in late February, but operation at normal capacity did not resume
until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
different policies regarding school starting at Easter, at which point rapid tests had
become widely available. We estimate the realized scenario to have essentially the same
effect as a situation with closed schools. Under fully opened schools with mandatory
tests, total infections would have been 7\% higher; this number rises to 20\% without
tests. These effect sizes are broadly in line with empirical studies, e.g.,
\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
another metric, the effective weekly reproduction number differs by $0.017$ and $0.045$,
respectively. In light of the large negative effects school closures have on children
and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
socio-economic status---these results in conjunction with hindsight bias suggest that
opening schools combined with a testing strategy would have been beneficial. In other
situations, and particular when rapid test are not available at scale, trade-offs may
well be different.

@@

Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
receiving tests, testing at the workplace has larger effects than mandating employees to
work from home. Whether the share of workers working at the usual workplace is reduced
or increased by ten percent hardly has an effect on infection rates. Making testing
mandatory twice a week---assuming independent compliance by employers and workers or
95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
employers to the level of March would have increased infections by the same amount.

Our analysis has shown that in the transition path to high levels of vaccination and
possibly thereafter, large-scale rapid testing can substitute for some NPIs. This comes
at a fraction of the cost. A week of the fairly strict lockdown in early 2021 is
estimated to have cost around 20~Euros per capita \citet{Wollmershauser2021}; retail
prices for rapid tests were below one Euro in early June 2021. Despite these large
effects, the results on testing likely understate the benefits. Disadvantaged groups are
less likely to be reached by testing campaigns relying on voluntary participation
\citep[e.g.][]{StillmanTonin2021}; at the same time they have a higher risk to contract
CoViD-19 \citep{KochInstitut2021a}. Mandatory tests at school and at the workplace will
extend more into these groups. The same goes for individuals who exhibit a low level of
compliance with CoViD-19-related regulations. Compared to vaccinations, rapid testing
programmes allow a much quicker roll-out, making it an  short-term alternative.
